---
title: "p8105_hw2_xs2447"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1
Load the Mr. Trash Wheel sheet
```{r}
mr_df = read_excel(
  "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N535"
  ) %>%
  janitor::clean_names() %>% 
  drop_na()
```

Load the 2018 Precipitation sheet
```{r}
prcp_2018_df = read_excel(
  "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "2018 Precipitation",
  range = "A2:B14"
) %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)
```

Load the 2019 Precipitation sheet
```{r}
prcp_2019_df = read_excel(
  "data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "2019 Precipitation",
  range = "A2:B14"
) %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)
```

Combine precipitation datasets
```{r}
prcp_df = 
  bind_rows(prcp_2018_df,prcp_2019_df) %>% 
  mutate(month = month.name[month])
```

The cleaned Mr.Trash Wheel dataset has 453 rows and 14 columns, including important variables such as dumspter number and sports balls. The cleaned 2018 Precipitation and 2019 Precipitation datasets each has 12 rows and 3 variables. Therefore, after combination, the precipitation dataset has 24 rows and 3 variables. The 3 variables are month in character, total precipitation in that month, and year number. The total precipitation in 2018 is `r sum(prcp_2018_df[,'total'])`. 

```{r}
sports_2019 = filter(mr_df, year == 2019)[,"sports_balls"]
```
The median number of sports balls in a dumpster in 2019 is `r apply(sports_2019, 2, median)`.

## Problem 2
Load and clean pols_month.csv
```{r}
pols_month_df = read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year","month","day"), "-") %>% 
  mutate(
    month = month.name[as.double(month)],
    month = tolower(substr(month, 1, 3)),
    president = if_else(prez_gop == 0, "dem","gop"),
    year = as.double(year)
    ) %>% 
  select(-prez_gop, -prez_dem, -day)
```

Load and clean snp.csv
```{r}
snp_df = read_csv("data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, c( "month","day","year"), "/") %>% 
  mutate(
    month = month.name[as.double(month)],
    year = as.double(year),
    year = if_else(year < 16, year + 2000, year + 1900),
    month = tolower(substr(month, 1, 3))
    ) %>% 
  select(-day) %>% 
  relocate(year)
```

Load and clean unemployment.csv
```{r}
unemployment_df = 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    jan:dec,
    names_to = "month",
    values_to = "unemployment_perc"
  )
```

Merge snp into pols
```{r}
pols_month_df = left_join(
  pols_month_df, snp_df, by = c("year", "month")
)
```

Merge unemployment into results
```{r}
result_df = left_join(
  pols_month_df, unemployment_df, by = c("year", "month")
)
```

The merged pols_month_df is the combination of pol_month.csv and snp.csv. It has 822 rows and 10 columns. The party that the president belongs to at each specific time is an important variable in this dataset. This varible ranges from year 1947 to 2015. Another important variable in this dataset is the closing values of the S&P stock index on the associated date. It contains the data range from year 1950 to 2015. In the result_df, the unemployment percentage is added. Therefore, it has 822 rows and 11 columns. The data of unemployment percentage ranges from 1948 to 2015.



